Changes in the justification of educational inequalities: The role of perceptions of inequality and meritocracy during the COVID pandemic

### 2024-03-26 Title-Page.Rmd
update title page

Word count:

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors recognize the support of ANID/FONDECYT grant 1210847 and of the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.

>>>>

Word count: 7663

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors recognize the support of the National Research and Development Agency (ANID) through the grant from the National Fund for Scientific and Technological Development (FONDECYT) N° 1210847; from the Center for Conflict and Social Cohesion Studies (COES) grant Financing Fund for Research Centers in Priority Areas (FONDAP) N° 1523A0005; and Millenium Nucleus on Digital Inequalities and Opportunities (NUDOS) NCS2022_046

### 2024-03-26 paper.Rmd
update title page

In this final part of the analysis, we contrast longitudinal hypotheses regarding changes in the relationship between meritocracy and justification of economic inequality over time. In hypothesis 5 we proposed that the association between the perception of meritocracy and inequality justification mitigates over time, as meritocratic ideals could have weakened due to critical situations associated with the COVID pandemic. We test this hypothesis through interaction effects, displayed in Table \@ref(tab:interact). Model 1 is shown as a baseline model, it is the same as Model 6 in Table 1 but it only displays the variables involved in the interaction for the sake of space (all other variables are controlled for). Model 2 adds the interaction between the perception of meritocracy based on talent and time (panel wave), whereas Model 3 does the same but now for the perception of meritocracy related to effort. As observed, only the effort variable shows a significant interaction with the time variable ($\beta$=0.01, p<.05), meaning that, on average, the association between the perception of meritocracy based on effort and the justification of inequality increases by 0.01 points in every measurement point.

>>>>

In this final part of the analysis, we contrast longitudinal hypotheses regarding changes in the relationship between meritocracy and justification of economic inequality over time. In hypothesis 5 we proposed that the association between the perception of meritocracy and inequality justification mitigates over time, as meritocratic ideals could have weakened due to critical situations associated with the COVID pandemic. We test this hypothesis through interaction effects, displayed in Table \@ref(tab:interact). Model 1 is shown as a baseline model, it is the same as Model 6 in Table \@ref(tab:multilevel) but it only displays the variables involved in the interaction for the sake of space (all other variables are controlled for). Model 2 adds the interaction between the perception of meritocracy based on talent and time (panel wave), whereas Model 3 does the same but now for the perception of meritocracy related to effort. As observed, only the effort variable shows a significant interaction with the time variable ($\beta$=0.01, p<.05), meaning that, on average, the association between the perception of meritocracy based on effort and the justification of inequality increases by 0.01 points in every measurement point.
